---
title: "Homework 4"
author: "Daniel Caley"
date: "7/10/2021"
output: 
  pdf_document:
  toc: TRUE
---

# Questions for this Week
The homework for Week 4 is exercises 7-10 on page 66

# Question 7
## Plant Growth - Summary
The built‐in `PlantGrowth` data set contains three different groups, each representing a different plant food diet (you may need to type `data(PlantGrowth)` to activate it). The group labeled `“ctrl”` is the control group, while the other two groups are each a different type of experimental treatment. Run the `summary()` command on `PlantGrowth` and explain the output. 

```{r}
data("PlantGrowth")

summary(PlantGrowth)

```


## Plant Growth - Histogram
Create a histogram of the ctrl group. 
As a hint about R syntax, here is one way that you can access the ctrl group data:
`PlantGrowth$weight[PlantGrowth$group==”ctrl”]`
Also create histograms of the `trt1` and `trt2` groups. 

```{r}

hist(PlantGrowth$weight[PlantGrowth$group=="ctrl"])
hist(PlantGrowth$weight[PlantGrowth$group=="trt1"])
hist(PlantGrowth$weight[PlantGrowth$group=="trt2"])

```

## Planet Growth - Discriptive Statistics
```{r}

sd(PlantGrowth$weight[PlantGrowth$group=="ctrl"])
sd(PlantGrowth$weight[PlantGrowth$group=="trt1"])
sd(PlantGrowth$weight[PlantGrowth$group=="trt2"])
mean(PlantGrowth$weight[PlantGrowth$group=="ctrl"])
mean(PlantGrowth$weight[PlantGrowth$group=="trt1"])
mean(PlantGrowth$weight[PlantGrowth$group=="trt2"])


```


## Interpreting Histogram
What can you say about the differences in the groups by looking at the histograms?

- The min and max for the ctrl is between 4.0 - 6.5, trl1 is between 3.5 - 6.5 without a 5.0 bin, and trl2 is between 5 and 6.
- Trl1 has 2 peaks where trl 2 only has 1 peak and the ctl has the highest peak.


\pagebreak

# Question 8
## PlantGrowth Boxplot
Create a `boxplot` of the plant growth data, using the model “weight ~ group.”

```{r}

boxplot(weight ~ group, PlantGrowth)


```

## Interpretting the Results
 What can you say about the differences in the groups by looking at the `boxplots` for the different groups?

- Each have a different median, the thick black line illustrates this.
- The the minimum value or the lower wisker is all different
- Although the maximum value or the upper wisker is different, each of them are not too far off.

\pagebreak
# Question 9
## T-Test ctrl vs. trt1
Run a t‐test to compare the means of `ctrl` and `trt1` in the `PlantGrowth` data. 


```{r}

dfCtrl <- PlantGrowth$weight[PlantGrowth$group=="ctrl"]
dfTrt1 <- PlantGrowth$weight[PlantGrowth$group=="trt1"]

t.test(dfCtrl, dfTrt1)

```

## T-Test ctrl vs. trt1 Interpreting the Results
Report and interpret the confidence interval. Make sure to include a carefully worded statement about what the confidence interval implies with respect to the population mean difference between the `ctrl` and `trt1` groups.

- 95% of the mean weight difference falls between -0.2875 and 1.0295.  Meaning that with 95% confidence the average weight difference between Ctrl and Trt1 will full between -0.2875 and 1.0295.  Although the p-value is 0.2504 which is quite a high p-value.

\pagebreak
# Question 10
## T-Test ctrl vs. trt2
Run a t‐test to compare the means of `ctrl` and `trt2` in the `PlantGrowth` data. 

```{r}

dfTrt2 <- PlantGrowth$weight[PlantGrowth$group=="trt2"]

t.test(dfCtrl, dfTrt2)

```


## T-Test ctrl vs. trt2 Interpreting the Results
Report and interpret the confidence interval.

- 95% of the mean weight difference falls between -0.9828 and 0.00512.  Meaning that with 95% confidence the average weight difference between Ctrl and Trt2 will full between -0.9828 and 0.00512.  Here though the p-value is 0.0479 which in statistics a p-value under 0.05 is pretty good. 

